Supporting Information:\\Conformational plasticity of phospholipid headgroups in simulations and experiments

### 2021-05-03 manuscriptPGPEsuppl.tex
Updates from Overleaf

\author{\small Am{\'e}lie Bacle}
\affiliation{\tiny Laboratoire Coop\'{e}ratif "Lipotoxicity and Channelopathies - ConicMeds", Universit\'{e} de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }

\author{Pavel Buslaev}
\affiliation{Nanoscience Center and Department of Chemistry, University of Jyväskylä, P.O. Box 35, 40014 Jyväskylä, Finland}
\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, 141701 Dolgoprudny, Russia}

@@

%      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\

@@

      POPC                   & CHARMM36 \cite{klauda10}            & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
      POPC:POPG (7:3)        & CHARMM36 \cite{klauda10,venable13}    & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}     & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}     & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}     & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}    & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}    & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}    & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\

@@

      POPC             & CHARMM36 \cite{klauda10}        &0          & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
      POPC:POPE (1:1)  & CHARMM36 \cite{klauda10,venable13}         & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\

@@

      POPC             & OPLS-MacRog \cite{rog16}     &0           & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPCmacrog300K} \\
      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0           & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1macrogT300K} \\

@@

      POPC             & Slipid \cite{jambeck12b}     &0          & 512 & 23943 & 0  &  298  & 170 & 100 & \cite{POPCslipid298K} \\
      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 128 & 5120  & 0  &  298  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\

@@

      POPC                   & GROMOS-CKP / ?? \cite{piggot12,??}  & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCgromosCKPT310K}  \\
      POPC:POPG (7:3)        & GROMOS-CKP / ?? \cite{piggot11,piggot12,??}  & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG3gromosCKPT310K} \\

@@

      POPC                   & Slipid \cite{jambeck12b}  & 0      & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
      POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90} & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\

@@

      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}       & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}    & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}    & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}   & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}   & 0.1& 150:150 & 31401 & 57 &  298  & 718 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}  & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\

@@

      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0     & 350:88 & 26265 & 0  &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020} \\
      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020100mMCaCl} \\
      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 1.0& 350:88 & 24840 & 475 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG80201000mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 1.0& 150:150 & 29865 & 569 &  298  & 600 & 400 & \cite{Lipid17eccPOPCPOPG50501000mMCaCl} \\

@@

      POPC             & Berger-POPC-07 \cite{ollila07a} & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
      POPC:POPE (1:1)  & Berger-POPE-04~\cite{devries04}  & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\

@@

    \todo{ion model for GROMOS-CKP?} \\ % TJP: Most likely standard GROMOS but not my simulations so I've left the ion reference

@@

\noindent {\it POPC:POPG mixture} \todo{Simulation details by A. Peon.}

@@

\noindent
{\it Josef Melcr}
%% Please write description of your contribution here

>>>>

\author{\scriptsize Am{\'e}lie Bacle}
\affiliation{Laboratoire Coop\'{e}ratif "Lipotoxicity and Channelopathies - ConicMeds", Universit\'{e} de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }

\author{\scriptsize Pavel Buslaev}
\affiliation{\tiny Nanoscience Center and Department of Chemistry, University of Jyväskylä, P.O. Box 35, 40014 Jyväskylä, Finland}
\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, 141701 Dolgoprudny, Russia}

@@

%      POPC                   & CHARMM36 \citenum{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\

@@

      POPC                   & CHARMM36 \cite{klauda10}            & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \citenum{POPCcharmmT310K}  \\
      POPC:POPG (7:3)        & CHARMM36 \cite{klauda10,venable13}    & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \citenum{POPC7POPG1charmm36}  \\
      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}     & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \citenum{CHARMM36POPCPOPG5050} \\ 
      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}     & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \citenum{CHARMM36POPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}     & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \citenum{CHARMM36POPCPOPG50501000mMCaCl} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}    & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \citenum{CHARMM36POPCPOPG4010} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}    & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \citenum{CHARMM36POPCPOPG4010100mMCaCl} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}    & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \citenum{CHARMM36POPCPOPG40101000mMCaCl} \\

@@

      POPC             & CHARMM36 \cite{klauda10}        &0          & 256 & 8704 & 0  &  300  & 300 & 250 & \citenum{POPCcharmm300K} \\
      POPC:POPE (1:1)  & CHARMM36 \cite{klauda10,venable13}         & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \citenum{POPC1POPE1charmm36} \\

@@

      POPC             & OPLS-MacRog \cite{rog16}     &0           & 128 & 5120 & 0  &  300  & 500 & 300 & \citenum{POPCmacrog300K} \\
      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0           & 128 & 5120 & 0  &  300  & 500 & 300 & \citenum{POPC1POPE1macrogT300K} \\

@@

      POPC             & Slipid \cite{jambeck12b}     &0          & 512 & 23943 & 0  &  298  & 170 & 100 & \citenum{POPCslipid298K} \\
      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 128 & 5120  & 0  &  298  & 500 & 300 & \citenum{POPC1POPE1slipidT298K} \\

@@

      POPC                   & GROMOS-CKP \cite{piggot12} & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \citenum{POPCgromosCKPT310K}  \\
      POPC:POPG (7:3)        & GROMOS-CKP / GROMOS54A7 \cite{piggot11,piggot12,schmid11} & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \citenum{POPC7POPG3gromosCKPT310K} \\

@@

      POPC                   & Slipid \cite{jambeck12b}  & 0      & 500     & 25000 & 0   &  310  & 500 & 100 & \citenum{POPCslipid301K}  \\
      POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90} & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \citenum{slipidPOPC70POPG30T310K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \citenum{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \citenum{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \citenum{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \citenum{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \citenum{slipidPOPC50POPG50T298K} \\

@@

      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}       & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \citenum{Lipid17POPCPOPG8020} \\
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}    & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \citenum{Lipid17POPCPOPG8020100mMCaCl} \\
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}    & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \citenum{Lipid17POPCPOPG80201000mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}   & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \citenum{Lipid17POPCPOPG5050} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}   & 0.1& 150:150 & 31401 & 57 &  298  & 718 & 198 & \citenum{Lipid17POPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}  & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \citenum{Lipid17POPCPOPG50501000mMCaCl} \\

@@

      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0     & 350:88 & 26265 & 0  &  298  & 400 & 300 & \citenum{Lipid17eccPOPCPOPG8020} \\
      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 300 & \citenum{Lipid17eccPOPCPOPG8020100mMCaCl} \\
      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 1.0& 350:88 & 24840 & 475 &  298  & 400 & 300 & \citenum{Lipid17eccPOPCPOPG80201000mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \citenum{Lipid17eccPOPCPOPG5050} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \citenum{Lipid17eccPOPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 1.0& 150:150 & 29865 & 569 &  298  & 600 & 400 & \citenum{Lipid17eccPOPCPOPG50501000mMCaCl} \\

@@

      POPC             & Berger-POPC-07 \cite{ollila07a} & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \citenum{POPCberger300K} \\
      POPC:POPE (1:1)  & Berger-POPE-04~\cite{devries04}  & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \citenum{POPC1POPE1berger} \\

@@

@@

\noindent {\it POPC:POPG mixture} 
Symmetric membranes of pure POPC and POPC:POPG 7:3 mixtures containing 500 lipids and 50 SPC water molecules per lipid were composed with CHARMM-GUI \cite{lee16}. The POPC:POPG 7:3 mixture contained also 150 sodium counter ions to balance the charge of PG lipids. GROMOS-CKP force field~\cite{piggot11,piggot12} was used for lipids and GROMOS54A7 \cite{schmid11} for the rest. Other simulation details are as in POPE and POPG systems.

@@

\noindent
{\it Josef Melcr}
created the analysis script \texttt{scripts/calcOrderParameters.py}, 
prepared the Lipid17ecc topology, 
and contributed to discussions and manuscript preparation.
